Discussion Draft

### 2020-03-25 doc/Discussion.rmd
Trimming down the discussion part

@@

A thing about this study that we were very interested in was examining the faster evolution in males with the high recombining rates.

BUT applying these above models to the traits connected to the faster male evolution is a bit tricker since none of these moedls were designed to explain within species variation in recombination rates. (most were designed to explain the general pattern of sexual dimorphism in meiotic recombination rates.)

@@

```{r evo.interference, include=FALSE}

#review results,

#why unexpected

#assuptions for predictions

#models that predict this pattern

@@

We had an unexpectedd result of correlated evolution of gwRR with interference strength. This is unexpected beacuse...

-The rapid evolution of the genome wide rate in our high rec strains is due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.

These results were unexcepted and differ ffrom the major patterns listed in the literature (otto payseur, argawal Korol. (it also goes against the logical assumptions models of fitting more crossovers along the SC chromosome axis space)

Yet (we are confident that this pattern is not due to errors).

 the IFD distributions show less variance in the high rec strains, which would have MORE Chromosomes (Chm1 to Chm7) compared to the low rec strains which have fewer 2Cos (presumably limited to large chromosomes).

**SACE**
a reversal in heterochiasmy direction would make sense (under the sae model) if it maintained 
this isn't an explict prediction, but accordding this this model -- stronger interference (such as we observed) would have the consequence of maintaining larger blocks of chromosomes together.

-Also:
	-These results are different from the sex differences because the difference is seen in both the raw and normalized IFD, indicateing the mechanism  / physcial strength (interference strength (signal) mediated through the SC has evolved.

	-The difference in interference patterns is driven by the difference in the rates of small IFD observations, (mean vs median). a pattern that is suggestive of selection agisnt short IFDs

There is a model that predicts this positive ocrrelation. The goldstein results (reviewed in ottoPayseur). when the number of crossovers (the obligate crossover assumption is more strict / the
breifly, the model simulates / sets up a scnario where a modifer acts through the probabilty of flanking regions having 0 or 2 crossovers. ith crossover assuance, recombination rate increases the further apart the loci are spaced (?)

@@

 1.Gamete selection

<the sex with the most variable reproductive fitness will be selected to have lower genome wide recombination rates>

From our results, this model would predict -- fit the results of:

this model doesn't predict that the heterochismy direction would reverse.

2. SACE

Extension of reduction principle. Results in large blocks of genetic areas being kept together in male meiosis.

-- this model is an extension of the modifer model presented in Lenormand

-- (basics)
 extends the lenomand focusing on two conditions. first that the stength of epistatsis in dependant on the cis or trans nature. the authors note that this would apply to gene archtechture (and increase with in cis ) -- where the regulatroy region and coding region of a gene evolve (together)

@@

 - the simulation present in the paper are too weak to support the rapid spread of such a modifer system (selection coefficients are too weak to support the evolution -- spread of such a modifer.)

- Other ways this model departs from our results,

-Always predicting male will be lower recombination rates (or risk breaking up the epistatsis blocks) -- which doesn't fit with our results of faster male specific evolution in gwRR.

However / while not explicitaly arameterized or predicted in this model,
 -- this model predicts larger chuncks of chromosomes will segregate together
 (as outlined in Veller et al)

@@

 3.Two-locus (protect for drivers)

This second modifer model was persented in [@brandvain2012scrambling] the basics of the model are that females will have a longer genetic map and more centromere biased recombination landscape compared to males.

this model would predict / fit the results of

 4. Segregation and physiological Spindle Differences

```{r Spindle.Model, include=FALSE}

#1. context of meiosis

#2. 
```

@@

LOOK at the oogenesis and spermatogenesis as the context for meiosis (and how )

lead us to ask, how the different contexts of meiosis could influence recombination rates (the recombinaiton landscape)

(not sex chromosome)

while there are many different forms, flavors of anisogamy (cite that one paper)

symetrical vs asymetrical gamete division

@@

for the distinct **evolutionary patterns** selection on the recombination landscape via the spindle (segregation checkpoint),

_leaky SAC_ for oocytes (is known) .. meaning that bivalents / homologs without proper tension force

(potentially this bad checkpoint is adaptive for thwarting centromere drive systems, which work through the SAC and tension signals)

how the SAC spindle works, correctly

evidence that the oocyte spindle is leaky

_precition that the leaky spindle makes relaxes selection_

aTMOC's  (because there are mulitple, these could be uneven across the cell)

_asymetrical division_

this makes egg weak to driving centromeres which (are segregation distortors), that means some centromeres segreation (rotate to the polar body side of the egg)

_number of placement of crossovers influencing spindle_

For this model we propose that the number and placement of crossovers -- will affect the

_strengths of model_

_weakness_

No very few spermatocyte spindle data  (it's assumed to )

_review the correct SAC_

References:
Schuh Ellenberg 2007 spindle

5. Model prediction
The models above which predict directional selection on the recombination landscape in males: SACE and Spindle differences are most parsimonious for the models review in this paper. The reduction principle wouldn't predict males diverge in genome wide rates and two locus modifier doesn't predict males gwRR would be higher than females.

@@

>>>>

- Keeping interference section breif (background, explaningillustrating spindle based selection model)

@@

A thing about this study that we were very interested in was examining the faster evolution in males with the high recombining rates. BUT applying these above models to the traits connected to the faster male evolution is a bit tricker since none of these moedls were designed to explain within species variation in recombination rates. (most were designed to explain the general pattern of sexual dimorphism in meiotic recombination rates.)

@@

```{r evo.interference, include=FALSE}
#review results,
#why unexpected, unexpected pattern is great for identifying most parimonious model / explanation
#(we're gonna spend alot of space on this section)

# more convincing that the pattern is real

#lil background, for interferences 
#assuptions for predictions (most modifer models, )

#models that predict this pattern (CONTROL FOR CO NUMBER -- PRIORITIZE THE WEIGHT /PROB of CO ASSURANCE)

#the pattern is driven in part by short IFD (when short IFD removed in reduced data set -- the low strain mean goes up, but they are still shrter)

@@

The strongest correlate of faster gwRR in males was intercrease interference strength. This is unexpected beacuse most literature and logical models assoicate more dense spacing of crossovers along chromosomes. Luckily having unexpected results is good for distingushing the most parimonius model. These results were unexcepted and differ ffrom the major patterns listed in the literature (otto payseur, argawal Korol. (it also goes against the logical assumptions models of fitting more crossovers along the SC chromosome axis space)

**The rapid evolution of the genome wide rate in our high rec strains is PARTLY due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.**

We considered carefully if this pattern was due to error -- and we are confident that this pattern is genuine -- goes aginst expected noise ALSO this is observed from other work in this lab -- and some empiracal studies (cattle) (plant bomblies).

<while these results don't clearly fit the chromatin organization and interference strength prediction... determining the clean pattern is beyond the scope of this paper>

HOW do the models above treat this unexpected correlation?

**typical modifer models -- are not good for predicting interferece (in mode that is relevant to these types of results) since the models don't parameterize for indiviual meiosis events.** We conclude that under the haploid selection and meiotic driver model any prediction for evolution of interference strength follows from the 'logical' expectation of a negative correlation.

**SACE**
While the SACE model doesn't explecitly make predictions aof the out come for interference strength (or averge spacing of crossovers, )
We note that the main predicted pattern under this model of maintaining larger blocks of chromosomes in linkage can asrise from stronger interference / greater spacing of the distance between two foci (also maximize rbar).

 extends the lenomand focusing on two conditions. first that the stength of epistatsis in dependant on the cis or trans nature. the authors note that this would apply to gene archtechture (and increase with in cis ) -- where the regulatroy region and coding region of a gene evolve (together)

Under the **COM** model -- the placement of crossovers (the recombination landscape) is generated by 'waves' / movement from the bouquet stage.  It is conceivable that the pattern of stronger interference could be generated from the waves/movements of chrmosomes in the early prophase stage this would come from a change in the movement or extension of the axis length.

**spindle based selection model**
There is a model that predicts this positive ocrrelation. The goldstein results (reviewed in ottoPayseur). when the number of crossovers (the obligate crossover assumption is more strict / the
breifly, the model simulates / sets up a scnario where a modifer acts through the probabilty of flanking regions having 0 or 2 crossovers. ith crossover assuance, recombination rate increases the further apart the loci are spaced (?)

@@

@@

 - the simulation present in the paper are too weak to support the rapid spread of such a modifer system (selection coefficients are too weak to support the evolution -- spread of such a modifer.)
- Other ways this model departs from our results, 
-Always predicting male will be lower recombination rates (or risk breaking up the epistatsis blocks) -- which doesn't fit with our results of faster male specific evolution in gwRR.
However / while not explicitaly arameterized or predicted in this model,
 -- this model predicts larger chuncks of chromosomes will segregate together
 (as outlined in Veller et al)

@@

# REVIEW

```{r final.review.outline, echo=FALSE}
# comes down to explaning the interference correlation

# spindel selection -- 
#good for sex differences because it frames heterochiasmy within the context of meiosis

#can fit the interference evolution due to the (pressumed) effect on bivalent shape

#caveats and unanswered parts

# SACE is runner up

@@

@@

@@

<5. Model prediction
The models above which predict directional selection on the recombination landscape in males: SACE and Spindle based selection  are most parsimonious for the models review in this paper. The reduction principle wouldn't predict males diverge in genome wide rates and two locus modifier doesn't predict males gwRR would be higher than females.>

<3.Two-locus (protect for drivers)
This second modifer model was persented in [@brandvain2012scrambling] the basics of the model are that females will have a longer genetic map and more centromere biased recombination landscape compared to males.

this model would predict / fit the results of>

<Also:
	-These results are different from the sex differences because the difference is seen in both the raw and normalized IFD, indicateing the mechanism  / physcial strength (interference strength (signal) mediated through the SC has evolved.>
